\documentclass[addpoints]{exam}

\usepackage[utf8]{inputenc}
\usepackage{float}
\usepackage{geometry}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[breaklinks]{hyperref}
\usepackage{listings}
\usepackage{fancyhdr}
\usepackage[english]{babel}
\usepackage{mdframed}
\usepackage{bookmark}
\usepackage{lipsum}
\usepackage{color}
\usepackage{psfrag}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage{pythonhighlight}
\usepackage{pagerange}
\usepackage{venndiagram}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{array}
\usepackage{url}
\usepackage{pagerange}
\usepackage[shortlabels]{enumitem}
\newtheorem{theorem}{Theorem}

% Header and footer.
\pagestyle{headandfoot}
\runningheadrule
\runningfootrule
\runningheader{MGMT 323}{Supply Chain Managment}{Activity 01}
% \runningfooter{}{Page \thepage\ of }{}
\runningfooter{}{\thepage \;of 3}{}
\firstpageheader{}{}{}

\boxedpoints
\printanswers
% \qformat{} %Comment this to number questions, uncomment this to not number questions

\newcommand\union\cup
\newcommand\inter\cap
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\olsi}[1]{\,\overline{\!{#1}}} % overline short italic

\title{Food for Thought \\ MGMT 323 - Supply Chain Management}
\author{Activity 01}
\date{Ali Muhammad Asad - aa07190}

\pgfplotsset{compat=1.18}

\begin{document}
\maketitle
\begin{sloppypar}
\begin{questions}
    \question \textbf{Why} K\&N has organized everything in house right from the beginning (acquiring eggs) all the way down to the final product placement at retail stores (raw chicken meat, frozen products etc.)?
    \begin{solution}
        K\&N prioitizes food safety, quality, and control on all stages of poultry production. By managing the entire supply chain from acquiring eggs to the final product placement, K\&N can ensure high standards of food quality and safety. This vertical integration proves to be quite an effective strategy for K\&N in the following ways:
        \begin{enumerate}
            \item \textbf{Quality Control} as mentioned above,
            \item \textbf{Cost Efficiency} as it allows K\&N to reduce costs by removing any middlemen such as suppliers and distributors. Thus, K\&N can control production more efficiently and offer competitive prices to customers,
            \item \textbf{Supply Chain Reliability and Market Leadership}. Managing the entire process in house gives K\&N more control over its supply chain, which in turn allows K\&N to be more responsive to market demands and changes. Thus it not only allows for a more reliable supply chain, but has also helped them become the market leader. 
            \item \textbf{Customer Relations}. The reliable supply chain, cost efficiency, market leadership, and quality control have all helped K\&N build strong customer relations. Customers trust K\&N for their quality products and services.
            \item \textbf{Innovation}. With complete control over their supply chain, K\&N has the flexibility to innovate and experiement on their product development.
        \end{enumerate}
    \end{solution}

    \question Can you identify some initiatives taken by K\&N to halt or prohibit waste generation?
    \begin{solution}
        There are several initiatives taken by K\&N to halt or prohibit waste generation. Some of them are as so:
        \begin{enumerate}
            \item \textbf{Environment-Controlled Poultry Houses:} The breeding process, and raising process of chickens is done in special environment controlled poultry houses to reduce risk of contamination, diseases and ensure better health of their chickens. Their houses are also equipped with automated heating, cooling and ventilation to ensure the optimum environment for the growth and health of their chicken.
            \item \textbf{Moniroting Health:} The health of the chickens is monitored throghout the breeding process to ensure quality of the chickens. The hatching eggs are also placed in sanitized incubators. Their feed also consists of natural grain products with vitamin supplements to ensure their ingredients and feed remain wholesome for better quality chickens and reduced chances of disease. 
            \item \textbf{Clean and Temperature Controlled Vehicles:} K\&N transports their chicks and goods in clean and temperature controlled vehicles to reduce risk of spoilage, disease, contamination. 
            \item \textbf{Nipple Drinkers:} The chickens are allowed to eat at will, to ensure they eat upto their fill, and drink fresh, pure water from nipple drinkers that release water at a controlled rate to avoid water contamination. 
        \end{enumerate}
    \end{solution}

    \question Can you find out some clues regarding the "Sustainability Initiatives" taken by K\&N? (Hint: Sustainability is comprised of three dimensions Economic (Profits), Social (People), Environmental (Planet).)
    \begin{solution}
        The following can be gathered regarding the "Sustainability Initiatives":

        \begin{enumerate}
            \item \textbf{Economic (Profits):} \begin{itemize}
                \item Individual Quick Freezing Technology: they use one of the best technologies in the world for quick freezing so that their products remain fresh for longer periods of time, while maintianing quality, thus optimizing resource consumptions which save them costs.
                \item Product Availability: their products are readily available at stores, retailers, and delivered to quality and food safety conscious restaurants, clubs, caterers, hotels etc
            \end{itemize}
            \item \textbf{Social (People):} \begin{itemize}
                \item Transparent supply chain: the transparent supply chain, with good quality food ensures a good relation with the customers.
                \item Halal Slaughtering Process in compliance with Shariah: this allows people to trust that the chicken is not only of good quality, but also religiously accepted. Their Halal certifications by various authentic instutitions further builds trust, authenticity and reliability.
                \item Veterinary Inspections and Lab Testing: this ensures the chickens remain good quality, and wholesome, thus improving customer satisfaction. 
                \item HACCP and FSSC certifications: ensures removal of diseases, and other contaminants
                \item No Artificial Preservatives, MSGs, Nitrates, or Nitrides for safe products
            \end{itemize}
            \item \textbf{Environmental (Planet):} \begin{itemize}
                \item Nipple Drinkers: the chickens are alloed to drink fresh, pure water from nipple drinkers that release water at a controlled rate to avoid water contamination and water borne diseases.
                \item Natural Grain: the feed for chickens consists of natural grain products with vitamin supplements to ensure their ingredients and feed remain wholesome for better quality chickens, and reducing need of resource-intensive feeds that damage the environment
            \end{itemize}
        \end{enumerate}
    \end{solution}

    \question Can you guess about the competitive priorities for K\&N? What are they?
    \begin{solution}
        \begin{itemize}
            \item \textbf{Quality}
            \item \textbf{Safety}
            \item \textbf{Sustainability}
            \item \textbf{Customer Satisfaction}
        \end{itemize}
    \end{solution}

\end{questions}
\end{sloppypar}
\end{document}

