\documentclass[a4paper]{exam}

\usepackage{amsmath,amssymb, amsthm}
\usepackage[a4paper]{geometry}
\usepackage{hyperref}
\usepackage{enumitem}

% \theoremstyle{theorem}
\newtheorem{theorem}{Theorem}

% \theoremstyle{claim}
\newtheorem{claim}{Claim}

\qformat{{\large\bf \thequestion. \thequestiontitle}\hfill}
\boxedpoints
\printanswers

\title{Weekly Challenge 14: Decidability}
\author{CS 212 Nature of Computation\\Habib University\\Ali Muhammad Asad - aa07190}
\date{Fall 2023}

\begin{document}
\maketitle

\begin{questions}

\titledquestion{Deciding Primes}

  Show that a language, $A$, is decidable iff $A \le_m PRIMES$, where $PRIMES$ is the set of all prime numbers.
  \begin{solution} \vspace*{-2mm}
    % A language $A$ is decidable if there exists a Turing Machine $M$ which halts on all inputs for any string $w$. 
    % A language $A$ is mapping reducible to a language $B$ if there exists a computable function $f$ such that for every string $w$, $w \in A$ iff $f(w) \in B$.
    \begin{enumerate}
      \item \textbf{If $A$ is decidable, then $A$ is mapping reducible to $PRIMES$} 
      
      Assume $A$ is decidable, then there exists a Turing Machine $M$ that decides $A$. We can construct a computable function $f$ such that for any string $w$, $ w \in A $ iff $f(w) \in PRIMES$. We define $f$ as follows: For each input $w$, compute $ M(w) $. If $ M(w) $ accepts, let $ f(w) $ be a fixed prime number. If $ M(w) $ rejects, let $ f(w) $ be a fixed composite number. Then this function $f$ is computable as $M$ is a decider, and thus halts on all inputs.
      
      Thus, $ w \in A \implies f(w) $ is prime, and $ w \in A \implies f(w) $ is composite. Therefore, $A$ is mapping reducible to $PRIMES$.

      \item \textbf{If $A$ is mapping reducible to $PRIMES$, then $A$ is decidable}
      
      Assume $A$ is mapping reducible to $PRIMES$. Then there exists a computable function $f$ such that for any string $w$, $ w \in A \text{ iff } f(w) \in PRIMES $.

      Since $PRIMES$ is decidable since there exists algorithms that can decide primes, we can construct a Turing Machine $PM$ that decides $A$ as follows: \vspace*{-2mm}
      \begin{enumerate}[label=(\roman*)]
        \item On input $w$, compute $f(w)$. \vspace*{-1mm}
        \item Run the algorithm that decides $PRIMES$ on $f(w)$. \vspace*{-1mm}
        \item If the algorithm accepts, accept. If the algorithm rejects, reject
      \end{enumerate}
      So the computability of $f$ ensures that $PM$ can always compute $f(w)$ and the decidability of $PRIMES$ ensures that $PM$ always halts. Thus, $PM$ is a decider for $A$. Therefore, $A$ is decidable.
    \end{enumerate}

    Hence, we can conclude that $A$ is decidable iff $A$ is mapping reducible to $PRIMES$. \vspace*{-4mm}
    \begin{flushright}
      $ \blacksquare $
    \end{flushright}
  \end{solution}
\end{questions}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
